---
title: "Hedge Fund Diversification with PCA"
author: "R. Morley"
date: "6 November 2018"
output: 
  html_document:
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(dplyr)
library(stringr)
library(tidyr)
library(xts)
library(ggcorrplot)
library(psych)
library(d3heatmap)
library(scales)
```

### Hedge Fund diversification - Eurekahedge fund of funds indices.  

What are the top-level diversification clues that can be gleaned from the hedge fund of funds indices complied by the data provider Eurekahedge (www.eurekahedge.com).

Download the Eurekahedge indices from the website:  
```{r}
# download the Eurekahedge indices
eurekahedge_url <- "http://www.eurekahedge.com/df/Eurekahedge_indices.zip"
download.file(eurekahedge_url, destfile = "EurekaHedge_Indices.zip", mode = 'wb')
unzip("EurekaHedge_Indices.zip")
eh       <- read.csv("EurekaHedge_Indices.csv", header = TRUE, stringsAsFactors = F)
eh$Date  <- as.Date(eh$Date, format = "%d-%b-%Y")
```


### Extract the 'fund of funds' indices.

Use __str_detect__ (stingr) combined with __filter__ (dplyr) to extract all indices with 'Fund of Funds Index' in the name.

Convert from long form to wide form using __spread__ (tidyr).

```{r}
# extract the 'fund of funds' indices

eh_FoF                       <- eh %>% filter(str_detect(Index.Name, pattern = "Fund of Funds Index"))  %>% 
                                group_by(EHIID, Index.Name) %>% 
                                select(Index.Name, EHIID) %>% 
                                summarise(Total = n()) %>% 
                                as.data.frame()

eh_FoF_EHIIDs                <- eh_FoF$EHIID %>% as.numeric()

eh_dataset_FoF_longform      <- eh %>% filter(EHIID %in% eh_FoF_EHIIDs) %>% select(Index.Name, Date, NAV)
eh_dataset_FoF_wide          <- spread(eh_dataset_FoF_longform, key = Index.Name, value = NAV)

ehColsFoF                    <- dim(eh_dataset_FoF_wide)[2]

# convert Date to date format
eh_dataset_FoF_wide$Date     <- as.Date(eh_dataset_FoF_wide$Date, format = "%Y-%m-%d")

# convert all FoF indicies to xts format
eh_dataset_FoF.xts           <- xts(eh_dataset_FoF_wide[,2:ehColsFoF]/100, order.by = eh_dataset_FoF_wide[,1])[-1]

# reduce the length of the names  by removing 'Eurekahedge' and 'Fund of Funds Index'
colnames(eh_dataset_FoF.xts) <- gsub(pattern = "Eurekahedge ", "", colnames(eh_dataset_FoF.xts))
colnames(eh_dataset_FoF.xts) <- gsub(pattern = " Fund of Funds Index", "", colnames(eh_dataset_FoF.xts))
```


###  Graphing the Correlations between the (fund of fund) indices.

The correlations show an aweful lot of red which indicates a high general correlation between strategies.

```{r}
ggcorrplot(cor(na.omit(eh_dataset_FoF.xts)), tl.cex = 8)
```

### Diversification using principal components analysis.

We take the fund of funds indices and group them together into separate investment __'bets'__ using Principal Components Analysis (principal from pysch).  This is the approach taken by the authors of the following texts:

  + Multi-Dimensional Diversification: Improving Portfolio Selection Using Principal Component Analysis by Dr Rufus G Rankin.
  
  + A Practitioner's Guide to Asset Allocation (Wiley Finance) by William Kinlaw, Mark P. Kritzman, David Turkington.


```{r}
pcaFoF = principal(cor(na.omit(eh_dataset_FoF.xts)), nfactors = 10, rotate = 'varimax')

pcaFoFvarPC1 <- pcaFoF$Vaccounted[2,1];pcaFoFvarPC2 <- pcaFoF$Vaccounted[2,2];pcaFoFvarPC3 <- pcaFoF$Vaccounted[2,3];pcaFoFvarPC4 <- pcaFoF$Vaccounted[2,4]
```

How many independent investment __'bets'__ explain at least 75% of the variation between the indices (the larger the better)

```{r}
pcaFoF$Vaccounted["Cumulative Var",] > 0.75
grep(TRUE, pcaFoF$Vaccounted["Cumulative Var",] > 0.75) %>% min()
```

If the principal components are *rotated* individual fund of funds indices are more clearly allocated to separately identifiable components or __investment 'bets'__ as shown by *dark blue* groupings.  

We see that the broadest, generic indices such as '(Global/North American) Relative Value' or 'Global Arbitrage' or 'Event Driven' all load on the same / first (rotated) principal component __bet__ (RC1) which explains c. `r percent(pcaFoFvarPC1)` of total variation between the indices.

The second __bet__ or RC4 is everything 'Emerging Market' and 'Asia Pacific' and explains an additional `r percent(pcaFoFvarPC2)` of variance.

The third orthagonal __bet__ is everthing 'Japanese' explaining `r percent(pcaFoFvarPC3)` of variance.

And the fourth independent __bet__ is the 'Macro' and 'CTA' strategies explaining a further `r percent(pcaFoFvarPC4)` of variance.

```{r}
pcaFoF$Structure[1:length(pcaFoF$values), 1:pcaFoF$factors] %>%
  d3heatmap(dendrogram = "row", colors = "Blues")
```

--


###  Summary

> In summary; from a top down view of the fund of funds indices the most effective way to __diversify__ is to 1) to diversify *regionally* such as Emerging Markets and Japan and 2) to incorporate Macro and CTAs into the mix.


